Inferring Invariants of Distributed Programs

### 2015-04-30 Doc/introback.tex

In this work we 
assume that state of each node is comprised of its variables. Not all variables of one node is affected by other nodes, we define the set of variables of one node whose values are affected by other nodes as distributed state variables of that node. We refer to set of distributed 
state variables of all nodes as distributed state of the system. We use daikon \cite{ernst2001dynamically} to infer invariants between variables, so we can infer linear ($y = ax + b$), ordering ($y < x$) and containment ($x \in y$) relationships that may hold between the distributed state variables of all nodes in the system.

@@

Invariants are properties that should hold in certain points in a program. Although previous works \cite{ernst2001dynamically} showed that invariant inference is feasible in sequential programs, inferring invariants in a distributed system poses unique challenges and still is an open problem. To infer invariants in a distributed system we should have global snapshots (log of variables) of the system. However, not every snapshot of the system is a valid one. Mandy and Lamport introduced the notion of cut and consistent cut as snapshot and valid snapshots of a system. In \cite{mattern1989virtual}, Mattern proposed an algorithm to find consistent cuts. The notion of distributed state first proposed by Ousterhout in \cite{ousterhout1991role} as ``information retained in one place that describes something, or is determined by something, somewhere else in the system", we assume that variables whose value is affected by other nodes consists a significant portion of distributed state of that node. We use dynamic data flow analysis to detect the distributed state of the system. We also use vector clocks to partially order snapshots of the nodes in our system and form consistent cuts.

Our approach is to automatically infer invariants that may exist in distributed state of our system. To this end, we execute distributed programs by running the test cases and dynamically select a set of variables in each node that likely represent the distributed state 
across our distributed nodes. We do so with dynamic data flow analysis of variables and selecting those that are affected by a send or receive operation. Finally we use Daikon to infer invariants among our distributed state variables. \\ \\
We can view our approach as an optimization. The naive approach to solve inference problem is to log all of the variables at each instruction execution and give all consistent cuts of the system to daikon to infer invariants between the variables. However, this naive approach is not efficient at all, if a total number of n instruction is executed during a run by all nodes in the system and on average we have m variables for each node, we should log $n*m$ variables but many of these variables are local to one node and are not affected by other nodes, so we may even end up finding many false positives (invariants that hold by accident and are not meaningful). Our approach only logs variables that an invariant might exist between them by using data flow analysis and finding group of variables (across the nodes) which affect each other. 
Another optimization that our approach does is that instead of logging the variables at each instruction, we only log them at consistent cuts annotated by developers. Based on our observation some must-hold invariants can be violated for small amount of time (e.g. execution time of several instructions) until systems goes to a stable state, so we decided to give developers responsibility to decide at which points in program invariants should hold and can be checked by annotating the code. By this approach we improve usefulness of our tool for developers. Another approach can be inferring the invariants from logs of the system, our approach removes the burden from developers to decide which portion of each node's state needs to be logged. We automatically infer variables that are likely to have invariants between them with data flow analysis.

>>>>

In this work we 
assume that state of each node is comprised of its variables. Not all variables of one node is affected by other nodes, we define the set of variables of one node whose values are affected by other nodes as distributed state variables of that node. We refer to set of distributed 
state variables of all nodes as distributed state of the system. We use Daikon \cite{ernst2001dynamically} to infer invariants between variables, so we can infer linear ($y = ax + b$), ordering ($y < x$) and containment ($x \in y$) relationships that may hold between the distributed state variables of all nodes in the system.

@@

Invariants are properties that should hold in certain points in a program. Although previous works \cite{ernst2001dynamically} showed that invariant inference is feasible in sequential programs, inferring invariants in a distributed system poses unique challenges and still is an open problem.To infer invariants in a distributed system we should have global snapshots (log of variables) of the system. However, not every snapshot of the system is a valid one. Mandy and Lamport introduced the notion of cut and consistent cut as snapshot and valid snapshots of a system. In \cite{mattern1989virtual}, Mattern proposed an algorithm to find consistent cuts. We used vector clocks \cite{vectorclock} to partially order snapshots of the nodes in our system and form consistent cuts. The notion of distributed state first proposed by Ousterhout in \cite{ousterhout1991role} as ``information retained in one place that describes something, or is determined by something, somewhere else in the system", we assume that variables whose value is affected by other nodes consists a significant portion of distributed state of that node. We use static data and control flow analysis to detect the distributed state of the system. To do so we computed the forward and backward slice of the program  \cite{programslice} from communication points in the program, such as receive and send functions, and determined list of variables whose values are affected by a receive function or can affect values that are sent by a send function. By doing so, we identified the variables that contribute to interactions between nodes, possibly affect values of variables in other nodes, hence form the distributed state of the program.

We can view our approach as an optimization. The naive approach to solve inference problem is to log all of the variables at each instruction execution and give all consistent cuts of the system to a invariant detector tool to infer invariants between the variables. However, this naive approach is not efficient at all, if a total number of $n$ instruction is executed during a run by all nodes in the system and on average we have $m$ variables for each node, we should log $n*m$ variables but many of these variables are local to one node and are not affected by other nodes, so we may even end up finding many false positives (invariants that hold by accident and are not meaningful). Our approach only logs variables that an invariant might exist between them by using data and control flow analysis and finding group of variables (across the nodes) which affect each other. 
Another optimization that our approach does is that instead of logging the variables at each instruction, we only log them at consistent cuts annotated by developers. Based on our observation some must-hold invariants can be violated for small amount of time (e.g., execution time of several instructions) until systems goes to a stable state, so we decided to give developers responsibility to decide at which points in program invariants should hold and can be checked, by annotating the code. By this approach we improve usefulness of our tool for developers. Another approach can be inferring the invariants from logs of the system, our approach removes the burden from developers to decide which portion of each node's state needs to be logged. We automatically infer the variables that are likely to have invariants between them with data and control flow analysis.
